Model Order Selection and Eigen Similarity based Framework for Detection and Identification of Network Attacks

### 2016-12-22 docs/papers/2016MONE/submission/respose01/mone-response_to_reviewers.txt

Response to the Handling Editor and Reviewers

Submission No.: MONE-D-16-00333
Title: Offline Mode for Corporate Mobile Client Security Architecture
SM 188 – Sustainable Computing Techniques for Mobile Networks
Mobile Networks and Applications

Dear Editor-in-Chief and reviewers,
The authors would like to thank the Associate Editor-in-Chief and the Reviewers for volunteering their time in reviewing our manuscript and providing us with valuable comments which allowed us to significantly improve the paper. We have carefully revised the manuscript based on the comments and suggestions. All changes in the paper have been marked with a color. The following is our point-by-point responses to the raised comments.

Kind regards,

The authors.

1. REVIEWER #1
1.1. COMMENT: The paper presents a reasonable approach to offline mode for reading secure documents on mobile clients without consuming much of the device resources under a restricted adversary model.
1.1.1. REPLY: Agradecer
1.2. COMMENT: The adversary model seems not complete. (1) The adversary model mentioned in Subsection 6.1 is not clear because affirmative statements are mixed with negative statements. (2) May the adversary disable the TIM?
1.2.1. REPLY: Aguardando resposta de Tanya
1.3. COMMENT: In Subsection 5.2, the description of ABE components is unclear: (1) typo of redundant double prime('') in PK, (2) no description of basic properties of bilinear paring function, (3) no description of the meanings of the notations w and s. (4) bad-form notation "for all i in {t_i}_M" (because "i" cannot belong to the set).
1.3.1. REPLY: Aguardando resposta de Tanya
1.4. COMMENT: In Subsection 5.3, the description of SSS (modular secret sharing) is unclear: (1) Di is a typo, (2) the notation based on Chinese Remainder Theory (CRT) is not well presented, (3) the meanings of d, S, p_0, and p_1 are not clear, (4) What are the reasons for S being equivalent to s and s_1 in modulo p_0 and p_1, respectively?
1.4.1. REPLY: Aguardando resposta de Tanya
1.5. COMMENT: In Subsection 6.2, it may be better to change "operations 1)-7)" to "actions 1)-7) (as mentioned in the offline mode)".
1.5.1. REPLY: Aguardando resposta de Tanya
1.6. COMMENT: There is little description about how to retrieve the logs from the Galaxy mobile devices and the correctness of anomaly detection in the experiments.
1.6.1. REPLY: Interagir com anton. Correctness vai ser tratado em trabalhos futuros

2. REVIEWER #3
2.1. COMMENT: In the related work, these following references need to be considered to enhance the completeness of the work. Heydari, Mohammad, et al. "An efficient password-based authenticated key exchange protocol with provable security for mobile client-client networks." Wireless Personal Communications 88.2 (2016). Mayrhofer, René. "An architecture for secure mobile devices." Security and Communication Networks 8.10 (2015): 1958-1970. Shila, Devu Manikantan, et al. "AMCloud: Towards a Secure Autonomic Mobile Ad Hoc Cloud Computing System." IEEE Wireless Communications(2016). Chang, Hyunseok, et al. "Design and Architecture of a Software Defined Proximity Cloud." Advances in Mobile Cloud Computing Systems (2015): 123. Xia, Yubin, et al. "TinMan: eliminating confidential mobile data exposure with security oriented offloading." Proceedings of the Tenth European Conference on Computer Systems. ACM, 2015. Kulkarni, Pallavi, and Rajashri Khanai. "Addressing mobile Cloud Computing security issues: a survey." Communications and Signal Processing (ICCSP), 2015 International Conference on. IEEE, 2015.
2.1.1. REPLY: Fazer
2.2. COMMENT: More discussions on the figure 1 and 2 should be added. The current version is not easy to understand.
2.2.1. REPLY: Tanya
2.3. COMMENT: The figure 8 should be explained more in detail.
2.3.1. REPLY: Tanya
2.4. COMMENT: The formula 5 should be discussed more, the related sentence should be introduced with more theoretical support. The set of obtained vectors y(q)m composes the zero mean matrix Y(q), then the zero mean covariance ma-trix ^R(q)=yy can be calculated as follows.
2.4.1. REPLY: Fazer

*** Definir data de retorno que seja suficiente para podermos fazer por conta própria se Tanya não responder, considerando o deadline de 11/01

### 2016-12-22 docs/papers/2016JNCA/2016JNCA.tex
preparacao para respostas aos revisores

\begin{abstract}
Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks. In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e. blind detection. This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis. In order to validate the proposed framework, we consider network traffic datasets that contains malicious activity such as flood and port probing attacks. We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity. The experiments performed in a real network and in a well known public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
\end{abstract}

@@

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations. Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}. Additionally, anomaly detection techniques can be categorized in classification, statistical, Information theory and clustering based \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.

In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks. Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). To the best of our knowledge there is no similar model in the literature. The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis. In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we also evaluate the accuracy and performance of the proposed framework applied to a simulated scenario and to the DARPA 1998 Dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset. Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.

@@

This paper is organized as follows. In Section \ref{sec:relatedworks}, related works are discussed. Section \ref{sec:datamodel} presents the data model and the evaluated datasets. Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks. Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discuss the computational complexity of the proposed framework and evaluates the required processing time for simulated scenarios. Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work. The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.

@@

Callegari \emph{et al} \cite{Zonglin2009} proposed a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focused in flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.

Lee \emph{et al.} \cite{Lee2013} proposed OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack. Additionally, Lee \emph{et al.} only evaluated their proposed scheme for covariance analysis, while we adopted an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.

@@

Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}. Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory. However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks. Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies. Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work was not applied to probe and flood attack detection, simultaneously.

@@

Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks. Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}. NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications. Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations. Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a simulated scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets. It is important to note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.

@@

This section presents details of the simulated scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic. Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks. Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition. Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discuss the use of the DARPA dataset for evaluation of the proposed approach.

@@

\begin{figure}[h!]
     \centering 
     \includegraphics[height=5cm, width=7cm]{results/figures/fig09.png}
     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
     \label{fig:fig1}
\end{figure}

@@

A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of transport protocol used. The following TCP traffic log is presented in order to exemplify the collected data:
\newline
\newline
\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
\newline

@@

The network attacks evaluated by this work are: synflood, fraggle and port scan. The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack. DoS attempts to block access to system or network resources, it is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks. Probe attacks scan the networks to to collect information about host, such as IP addresses, ports and services.

@@

The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks. The attacks in this dataset
can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
remote machine; user to root (U2R), which is characterized by unauthorized access to local
super-user privileges; and probe attack. Since the proposed approach focus on flood and probe attack, we analyze the attacks that present behaviors similar to flood or probe attack. We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack. One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior. Therefore, we select the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets. From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.

@@

If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute. Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack was identified, according to

@@

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the simulated scenario adopted in the experiments is summarized. Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the simulated scenario. In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset. Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the simulated scenario. Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.

\subsection{Simulated Scenario}
\label{sec:AnalyzedScenario}

@@

This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$. Hence, the proposed framework was applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.

@@

Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches. For the incremental individualized approach, each minute was incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack. Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated. For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis. For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.

@@

This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset. Since the proposed framework is detailed explained in previous sections, here we focus on the parameter selection, dataset evaluation and results for flood and probe attack identification.

The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data. The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack. For this evaluation, we performed an evaluation per day, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregated by minute and by port number. For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ was obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.

Since the proposed framework focus on flood and probe attack detection, we evaluated only the attacks with behavior similar to flood or probe attack. Initially we selected all DoS and probe attacks, but we observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning. Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors. We select the following week-day-attack cases:

@@

The analysis based on sample covariation of zero mean variables was evaluated for flooding behavior of netpune attacks, obtaining rates of 100\% for true positive (TP) detection and 60\% for false positive (FP) detection from 30 time frames. The results also show 50\% of misclassification rate \cite{bhuyan2014network}, which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed. 
The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port. This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.

The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability. The results show rates of 76.9\% for TP detection and 23\% for FP detection from 94 time frames. The observed misclassification rate for this scenario was 32.7\%. We observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks. Cases with a delay of one minute or more were not detected by the proposed approach.

The performance of detection rate of flooding attacks is compared with the method proposed by Callegari \emph{et al} \cite{Zonglin2009}. This work is a statistical method, based on PCA, without training or learning methods, even though it relies on visual analysis for principal components selection. The best detection rate of \cite{Zonglin2009} was 82\% for detection of synthetically added flood attacks, while our proposal obtains 100\% of detection rate for detection of flood attacks of DARPA dataset.

Due to the lack of statistical techniques without training or learning methods for detection of probe attacks, we compare our approach with Lu and Ghorbani's \cite{Lu2009} proposal, which is a network anomaly detection model based on signal processing techniques that uses DARPA dataset for evaluation. The results of \cite{Lu2009} show the best detection rate of 94.67\% in terms of general attack instance detection, but shows a case case with 50\% of attack instance detection for the portsweep attack. The proposed approach presents 76.9\% of detection rate measured specifically for probe attacks, without the requirement of learning or training methods, in contrast to Lu and Ghorbani's \cite{Lu2009} work.

\section{Complexity Analysis}
\label{sec:Complexity}

This section discusses the computational complexity of the proposed framework, focusing on the main steps, which are the eigenvalues decomposition (EVD), largest eigenvalues analysis, application of MOS scheme and eigen similarity analysis, according to Figure \ref{fig:fig80} and equations presented in Section \ref{sec:prop_getv}.

The EVD, calculated according to (\ref{eq:eq06}), requires the previous calculation of covariance matrix, according to Equations \ref{eq:eq02}, \ref{eq:eq03}, \ref{eq:eq04} and \ref{eq:eq05}. The covariance matrix calculation is $O(M^2N)$ and the EVD is $O(N^3)$, where $M$ denotes the number of network ports and $N$ denotes the period time. Therefore, the computational complexity for all steps for EVD can be represented as $O(M^2N + N^3)$ and yields an $O(N^3)$ upper bound on the worst-case running time for EVD.

>>>>

\begin{abstract}
Novel schemes for attack detection are crucial to identify adaptive malicious traffic coming from sources that are quickly mobilized by attackers in high throughput communication networks. In this context, signal processing techniques have been applied to attack detection due to their capability to detect anomalies that are previously unknown, i.e. blind detection. This paper proposes a signal processing framework for the detection and identification of network attacks using concepts of model order selection (MOS), eigenvalues and similarity analysis. In order to validate the proposed framework, we consider network traffic datasets that contain malicious activity such as flood and port probing attacks. We propose to model the network traffic as a superposition of components, namely, user's operations (legitimate traffic), network service operation not related to the user (noise) and the malicious activity. The experiments performed in a real network and also using the DARPA 1998 public dataset show that the proposed blind detection approach achieves satisfactory levels of accuracy in terms of timely detection and identification of TCP/UDP ports under attack.
\end{abstract}

@@

Information security may consist of both technical and procedural aspects. The former includes equipment and security systems, while the latter corresponds to security rules and recommendations. Intrusion detection and intrusion prevention systems are security systems used, respectively, to detect (passively) and prevent (proactively) threats to computer systems and computer networks. Such systems can work in the following fashions: signature-based, anomaly-based or hybrid \cite{Huang2009,mudzingwa2012study}. Additionally, anomaly detection techniques can be categorized in classification, statistical, information theory and clustering based, according to \cite{bhuyan2014network,ahmed2016survey,osanaiye2016distributed}.

In the context of anomaly-based schemes, this work proposes a statistical approach based on signal processing techniques for the detection of malicious traffic in computer networks. Inspired by \cite{david2011blind,da2012improved}, this work models the network traffic using a signal processing formulation as a composition of three components: legitimate traffic, malicious traffic and noise, taking into account the incoming and outgoing traffic in certain types of network ports (TCP or UDP). To the best of our knowledge there is no similar model in the literature. The proposed technique is based on eigenvalue analysis, model order selection (MOS) and similarity analysis. In contrast to \cite{david2011blind,da2012improved,tenorio2013greatest}, MOS and eigenvalue analysis are applied to detect time frames under attack. In addition, we also evaluate the accuracy and performance of the proposed framework applied to a experimental scenario and to the DARPA 1998 dataset \citep{osanaiye2016distributed}, which is a well known network traffic dataset. Furthermore, this proposed approach has its accuracy evaluation based on eigen similarity analysis for extracting detailed information about accurate time and network ports under attack.

@@

This paper is organized as follows. In Section \ref{sec:relatedworks}, related works are discussed. Section \ref{sec:datamodel} presents the data model and the evaluated datasets. Section \ref{sec:prop_getv} describes the proposed framework for blind and automatic detection of flood and probe attacks. Section \ref{sec:experimentalresults} discusses the experimental validation and presents the results, and Section \ref{sec:Complexity} discusses the computational complexity of the proposed framework and evaluates the required processing time for tested scenarios. Section \ref{sec:conclusionandfutureworks} draws the conclusions and the suggestions for future work. The \ref{sec:mos} presents mathematical concepts of examples of state-of-the-art MOS schemes.

@@

Callegari \emph{et al} \cite{Zonglin2009} propose a PCA-based method for identifying the traffic flows responsible for an anomaly detected at the aggregate level and evaluated their proposal through a dataset with synthetic anomalies added in the data. However, Callegari \emph{et al} focus on flood attack detection, not addressing probe attack detection, and their approach relies on visual analysis.

Lee \emph{et al.} \cite{Lee2013} propose OverSampling PCA (osPCA), which allows one to determine the anomaly of the target instance according to the variation of the resulting dominant eigenvector obtained by similarity analysis and over sampling. In contrast to Lee \emph{et al.}, the framework applies MOS for detection of time frames under attack and similarity analysis to extract details for detection of time and ports under attack. Additionally, Lee \emph{et al.} only evaluate their proposed scheme for covariance analysis, while we adopt an analysis based on sample covariance of zero mean variables and sample covariance of zero mean and unitary standard deviation variables, for flood and probe attacks, respectively.

@@

Signal processing techniques have been successfully applied to network anomaly detection \cite{Lu2009}. Lu and Ghorbani \cite{Lu2009} proposed a network anomaly detection model based on network flow, wavelet approximation, and system identification theory. However, their work requires a training method to produce a prediction model for normal daily traffic and presents limitations on identification of behaviors without significant outliers, such as port scan attacks. Zonglin \emph{et al} \cite{Zonglin2009} proposed a method to detect traffic anomaly with correlation analysis, where the correlation between traffic signals and the predicted traffic signals are used to reveal anomalies. Zonglin \emph{et al.} \cite{Zonglin2009} evaluated the correlation analysis for anomaly detection, but the work is not applied to probe and flood attack detection, simultaneously.

@@

Several approaches for network attack detections uses the KDD 99 \cite{ji2016multi,ahmed2016survey,osanaiye2016distributed,bhuyan2014network} datasets for accuracy and performance evaluation, due to their availability and labeled attacks. Even though the KDD 99 dataset are criticized by for the generation procedure and the risk of over-estimations of anomaly detection due to data redundancy, it still represents one of the few publicly available labeled datasets currently in use today by researchers \cite{osanaiye2016distributed,bhuyan2014network}. NSL-KDD \cite{tavallaee2009detailed} dataset is the refined version of the KDD 99 dataset that redundant data records are removed, in order to avoid biased classifications. Additionally, some approaches uses simulated \cite{callegari2011novel} scenarios or non-public datasets their evaluations. Since the proposed approach relies on a packet level analysis and the KDD 99 and NSL-KDD datasets adopt a traffic aggregation by connections, we consider the use of a experimental scenario on a real network and the DARPA 1998 dataset, which is the source for the creation of the KDD 99 and NSL-KDD datasets. Note that the proposed approach is not based on learning or classification techniques, which are more susceptible to biased results caused by the issues in the DARPA/KDD datasets.

@@

This section presents details of the experimental scenario and the selected cases of the DARPA dataset, along with a description of the dataset model as a signal superposition of legitimate traffic, noise and malicious traffic. Subsection \ref{sec:DataCollection} describes the environment and scenario adopted in order to reproduce flood and probe attacks. Subsection \ref{sec:ModelingData} presents how network traffic can be modeled as signal superposition. Subsection \ref{sec:SynfloodFraggleandPortscan} details the traffic of synflood, fraggle and port scan attacks, and Subsection \ref{sec:DarpaDataset} discusses the use of the DARPA dataset for evaluation of the proposed approach.

@@

\begin{figure}[h!]
     \centering 
     \includegraphics[height=6cm, width=9cm]{results/figures/fig09.png}
     \caption{Scenario to reproduce legitimate traffic, noise, flood and port scan.}
     \label{fig:fig1}
\end{figure}

@@

A network traffic log is commonly formed by timestamp, protocol, source IP address, source port, destination IP address, destination port and additional information, according to the type of the used transport protocol. The following TCP traffic log is presented in order to exemplify the collected data:
\newline
\newline
\texttt{\justify21:00:34.099289 IP 192.168.1.102.34712 > 200.221.2.45.80: Flags [S], seq 2424058224, win 14600, options [mss 1460, sackOK,TS val 244136 ecr 0,nop,wscale 7], length 0}
\newline

@@

The network attacks evaluated by this work are: synflood, fraggle and port scan. The first two attacks can be qualified as flood or denial of service (DoS) attacks, while the last one can be qualified as probe or port scanning attack. DoS attempts to block access to system or network resources and is implemented by either forcing targets to be unavailable through the exploiting of system vulnerabilities, or consuming resources through large amount of network traffic, characterizing flood attacks. Probe attacks scan the networks to collect information about host, such as IP addresses, ports and services.

@@

The DARPA 1998 dataset\footnote{https://www.ll.mit.edu/ideval/data/} includes 7 weeks of sniffed traffic saved into raw TCPDUMP packet data, from inside and outside origins, with labeled attacks. The attacks in this dataset
can be grouped into: denial-of-service (DoS); remote to local (R2L), which is characterized by unauthorized access from a
remote machine; user to root (U2R), which is characterized by unauthorized access to local
super-user privileges; and probe attack. Since the proposed approach focus on flood and probe attack, the analysis concentrates on the attacks that present behaviors similar to flood or probe attack. We observe that the most cases of DoS focus on exploit system vulnerabilities instead of on flooding attack. One example is the occurrence of a neptune attack which sends 20 only SYN packets, what is a behavior that differs of the expected flood attack behavior. Therefore, there were selected the cases that simulates several network traffic or numerous connection requests, also known as flooding attack \cite{ahmed2016survey,osanaiye2016distributed}, and the cases that scan ports sending just a few packets. From the simulated probe attacks, we select the cases that rely on TCP or UDP connections.

@@

If $s_n = 1$, then the two eigenvectors are completely similar and no anomaly is detected. Smaller values of $s_n$ mean less similarity and can indicate an anomaly, according to a defined 
threshold $l$, assuming that if $s_n < l$, then a network attack is identified during the $n$-th minute. Therefore, the $s_n$ of each $n$-th minute shall be compared with the threshold $l$ to evaluate if a attack is identified, according to

@@

This section presents the performed experiments and the acquired results. First, in Section \ref{sec:AnalyzedScenario}, the experimental scenario adopted in the experiments is summarized. Then, Section \ref{sec:largesteigenvaluesanalysis} shows the results of the largest eigenvalue analysis by time frames for the experimental scenario. In Section \ref{sec:MOSSchemesEvaluation} are described the results of the evaluated MOS schemes for attack detection in the simulated dataset. Section \ref{sec:EigenvalueAnalysis} presents the results of the eigenvalue analysis for identification of time frames under attack, Section \ref{sec:EigenSimilarityAnalysis} shows the results of similarity analysis for detailed flood and port scan identification for the experimental scenario. Section \ref{sec:DarpaEvaluation} presents the results of the largest eigenvalue analysis, model order selection and the eigenvalue analysis for flood and probe attack detection in the DARPA 1998 dataset.

\subsection{Experimental Scenario}
\label{sec:AnalyzedScenario}

@@

This paper proposes applying eigen similarity analysis to detect time and ports under attack, from each $q$-th time frames under attack defined by $\rm{\boldsymbol{\hat{q}}}_{\rm max}$. Hence, the proposed framework is applied to the time frames where $q=3$, $q=4$ and $q=5$ to respectively evaluate its effectiveness for port scan, synflood and fraggle attack detection.

@@

Three approaches were evaluated for eigen similarity analysis: incremental, individual and incremental individualized approaches. For the incremental individualized approach, each minute is incrementally appended into the selected $\boldsymbol{X}^{(q)}$ for obtaining $\boldsymbol{v}_{(n)}$ to similarity analysis of the $n$-th minute, until detect the first $n$-th minute under attack. Subsequently, $\boldsymbol{X}_n$ became the new reference of traffic without network attack and each subsequent minute must have its similarity individually evaluated. For the incremental approach, each $n$-th minute must be incrementally appended into $\boldsymbol{X}^{(q)}$, for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual time similarity analyis. For the individual approach, each $n$-th minute must be individually appended into $\boldsymbol{X}^{(q)}$, without incremental append, but doing individual appended into $\boldsymbol{X}^{(q)}$ for obtaining the next eigenvectors $\boldsymbol{v}_{(n)}$ for individual similarity analyis.

@@

This subsection presents a summarized view of results obtained from the application of the proposed framework, focusing on the largest eigenvalue analysis, model order selection and the eigenvalue analysis, for flood and probe attack detection in the DARPA 1998 dataset. Since the proposed framework is detailed in Section \ref{sec:prop_getv} and in Subsections \ref{sec:AnalyzedScenario}, \ref{sec:largesteigenvaluesanalysis}, \ref{sec:MOSSchemesEvaluation}, \ref{sec:EigenvalueAnalysis} and \ref{sec:EigenSimilarityAnalysis}, here the focus is on the parameter selection, dataset evaluation and results for flood and probe attack identification.

The DARPA dataset includes 7 weeks of sniffed traffic saved into raw packet data. The traffic and the labeled attacks are grouped by week and day, with information of the number and types of attacks per day, but also providing the start time for each labeled attack. For this evaluation, an evaluation per day was performed, considering the network traffic of 24 hours split into $Q$ time frames of 60 minutes ($N = 60$) and aggregate by minute and by port number. For each time frame $q$, a traffic matrix $\boldsymbol{X}^{(q)} \in \mathbb{R}^{17 \times 20}$ is obtained, considering the ports 20, 21, 22, 23, 25, 79, 80, 88, 107, 109, 110, 113, 115, 143, 161, 389 and 443.

Since the proposed framework focus on flood and probe attack detection, only the attacks with behavior similar to flood or probe attack were evaluated. Initially all DoS and probe attacks were selected, but it was observed that the most cases of DoS focus on exploit system vulnerabilities instead of flooding attack, and most of probe attacks focus on ICMP intead of port scanning. Therefore, for evaluation of the proposed approach for flood and probe attack detection, it is necessary to select cases that implements flood or port scan behaviors. The following week-day-attack cases  were selected:

@@

\begin{table*}[!t]
	\caption{Results of the attack detection evaluation}
	\label{tab:tab12}
	\centering
	\begin{tabular}{|c|c|c|c|c|}
		\hline \rowcolor{Gray} \begin{tabular}[x]{@{}l@{}}Solution\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Attack Type\end{tabular}	 & \begin{tabular}[x]{@{}l@{}}Metric\end{tabular}	& \begin{tabular}[x]{@{}l@{}}Result\end{tabular} \\ \hline
	Proposed Work	&Flooding	&True Positive	&100.00 \%\\ \hline
	Proposed Work	&Flooding	&False Positive	&60.00 \%\\ \hline
	Proposed Work	&Flooding	&Misclassification	&50.00 \%\\ \hline
	Proposed Work	&Probe	&True Positive	&76.92 \%\\ \hline
	Proposed Work	&Probe	&False Positive	&18.52 \%\\ \hline
	Proposed Work	&Probe	&Misclassification	&32.73 \%\\ \hline
	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&True Positive	&82.00 \%\\ \hline
	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&False Positive	&-\\ \hline
	Callegari \emph{et al} \cite{Zonglin2009}	&Flooding	&Misclassification	&-\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Overall	&True Positive	&94.67 \%\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Overall	&False Positive	&-\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Overall	&Misclassification	&-\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&True Positive	&50.00 \%\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&False Positive	&-\\ \hline
	Lu and Ghorbani \cite{Lu2009}	&Portsweep	&Misclassification	&-\\ \hline
	\end{tabular}
\end{table*}

Table \ref{tab:tab12} presents the evaluated results for attack detection, considering rates of TP, FP \cite{fleiss2013statistical} and misclassification, which is defined as $\frac{(FN+FP)}{(TP+FP+FN+TN)}$ \cite{bhuyan2014network}. The analysis based on sample covariation of zero mean variables is evaluated for flooding behavior of netpune attacks, obtaining rates of 100.00 \% for true positive (TP) detection and 60.00 \% for false positive (FP) detection from 30 time frames. The results also show 50.00 \% of misclassification rate , which attempts to estimate the probability of disagreement between the true and predicted cases by dividing the sum of FN and FP by the total number of pairs observed. 
The result for FP and misclassification analysis is poor due to the legitimate traffic of DARPA dataset presents high number of packets per time  from one source to one target, with no variation on IP source or target port. This observation corroborates with previous evaluations of the DARPA dataset that highlight issues regarding traffic redundancy.

The analysis based on covariance of zero mean and unitary standard deviation variables was evaluated for port scan attacks, including probe attacks and DoS attacks that send few packets for several ports in order to exploit some vulnerability. The results show rates of 76.92 \% for TP detection and 18.52 \% for FP detection from 94 time frames. The observed misclassification rate for this scenario is 32.73 \%. It was observed that all FN cases are probe attacks with a time delay between scanning one port and start scanning the next port, what can be called as sparse probe attacks. Cases with a delay of one minute or more were not detected by the proposed approach.
